Stateless Model Checking with Data-Race Preemption Points

### 2015-11-17 pldi-16/evaluation.tex
minor rephrasing

{\bf Finding the same bugs faster.}
To show that Iterative Deepening is effective even for MC domains without data races, such as message-passing distributed systems,
we ran the test suite with \quicksand~configured to explore only subsets of the hard-coded mutex PPs
(i.e., ignoring all data-race candidates).
%\footnote{
%Because \quicksand~is not yet instrumented to subset hard-coded PPs beyond the 4 ways shown in Figure~\ref{fig:id},
%we ran these tests for 2.5 hours on 4 CPUs each.
%Future work could parallelize QS-no-DR-PPs further; see \sect{\ref{sec:future}}.}.
%
The line QS-no-DR-PPs represents this experiment.
Even though SSS-MC mostly catches up to it by the end of the 10-hour budget,
QS-no-DR-PPs finds many more of the bugs much sooner.
Hence, for smaller arbitrary CPU budgets,
Iterative Deepening is likely to find bugs SSS-MC will miss.
%and for more ambitiously-sized tests,
%programmers can be more confident in the verification provided when \quicksand~times out with no bug found.
%We conclude that for smaller arbitrary CPU budgets, especially less than 1 hour,
%Iterative Deepening is likely to find bugs SSS-MC will miss.
%Moreover, it is also easy to imagine scaling up the size of each test case to test ,
%using more threads or longer sequences of API calls.
%We hope that is compelling even to users willing to spend many CPU-hours on testing.
%These results show explicitly that for arbitrary CPU-time budgets

@@

{\bf Finding new data-race bugs.}
The line QS-DRs represents our ``no holds barred'' \quicksand~tests,
with dynamically-discovered data-race PPs enabled:
we quickly pull ahead of SSS-MC, and ultimately conclude with 179\% as many bugs in total.
The break-even point is at a negligible 90 seconds.

@@

\begin{table*}[t]
	\begin{center}
		%\small
	\begin{tabular}{r|c||c|c|c||c|c|c|c|c}
		% TODO: Make the "bugs" left half of this graph into a bar graph.
		& {\bf Num.} & {\bf SSS-MC} & {\bf QS} & {\bf DR-only} & {\bf Total} & {\bf Untested} & {\bf Malloc-} & {\bf Nondet.} & {\bf Nondet.} \\
		{\bf Test name} & {\bf tested} & {\bf bugs} & {\bf bugs} & {\bf bugs} & {\bf DR PPs} & {\bf DR PPs} & {\bf recycle DRs} & {\bf DR PPs} & {\bf DR bugs} \\
		\hline
		{\bct		} & 79	& 7	& 11	& 4	&	&	& 65	&	& 4	\\
		{\tej		} & 79	& 14 	& 25	& 11	&	&	& 471	&	& 6	\\
		{\mxtest	} & 79	& 1	& 10	& 9	&	&	& 7	&	& 1	\\
		{\paradise	} & 79	& 10	& 17	& 7	&	&	& 268	&	& 3	\\
		{\paraguay	} & 79	& 4	& 12	& 9	&	&	& 188	&	& 0	\\
		{\rwldgr	} & 79	& 26	& 30	& 3	&	&	& 157	&	& 3	\\
		\hline
		{\prisema	} & 59	&	&	&	&	&	& 0	&	&	\\
		{\alarmsimul	} & 44	&	&	&	&	&	& 37	&	&	\\
		{\waitsimple	} & 52	&	&	&	&	&	& 57	&	&	\\
		\hline
		{\bf Total}	& 629	&	&	&	&	&	& 1250	&	&
	\end{tabular}
	\end{center}
	\caption{Summary of number and types of bugs found by each test program.
		SSS-MC is the control; QS is \quicksand.
		``DR-only'' counts among \quicksand's bugs how many required data-race PPs to expose (\sect{\ref{sec:eval-sssmc}}).
	``Nondeterministic DRs'' counts among the DR-only bugs how many candidates required MC integration to identify (\sect{\ref{sec:eval-falseneg}}).
		TODO: more explanation}
	\label{tab:drbugs}
\end{table*}

%\begin{figure}[t]
%	\includegraphics[width=0.48\textwidth]{bargraph.pdf}
%	\caption{{\bf TODO:} Is this graph productive to include in the paper? Or should I just leave the numbers in table format (the left half of Table~\ref{tab:drbugs})?}
%	\label{fig:bargraph}
%\end{figure}

%Figure~\ref{fig:bargraph}
The left half of Table~\ref{tab:drbugs}
breaks down the number and types of bugs found by each test program.
% The astute observer will note that the sum of all types of failures observed is more than the total number of data-race bugs...
%TODO CAMREADY: Run a mutex expt where "all atomic instrs" are PPs. See how many bugs are missed anyway.
% Wwe re-ran the \mxtest control experiment with \landslide~hard-coded to preempt on any atomic instruction
% (as well as on the mutex API boundaries).
% Still, this smarter configuration for SSS-MC found only 99999999999 bugs of \quicksand's 13.
In \mxtest~in particular, the control experiment found dramatically fewer bugs (just 1),
even compared to the other test programs\footnote{
	The one bug SSS-MC found was in a fully-assembly lock implementation. {\tt yield()}'s return value clobbered a value stored in {\tt \%eax}, which could lead to a failure after two repeated contentions. Preempting only on {\tt yield()} (in the contention loop) was sufficient.}.
%Intuitively, this is due to our control experiment being able to preempt only on the boundaries of the API which
%Though for many applications of MC, assuming a correct lock implementation is sufficient,
Though it suffices for many applications of MC to assume correctly-implemented locks,
we consider this strong evidence that low-level synchronization code must be verified with data-race PPs.

@@

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{nondet-drs-1-v2.pdf}
	\caption{Some data-race candidates may not be identified during a single program execution.
		Using nondeterministic data races as PPs,
		\quicksand~found 189\% as many bugs as could be found using single-pass data races alone.
	}
	\label{fig:dr-falsenegs}
\end{figure}
Figure~\ref{fig:dr-falsenegs} compares the types of data-race candidates necessary to expose each data-race bug in our test suite.
The first series represents the bugs found using PPs from single-pass data-race candidates,
% TODO: re read racefuzzer/portend paper to confirm this is true.
i.e., the state-of-the-art approach used by \cite{racefuzzer,portend}.
The second series shows all data-race bugs \quicksand~found,
which includes the former type as well as new bugs involving nondeterministic data-races.
\quicksand~found 68 data-race bugs in total, only 36 of which could be found with single-pass data-race candidates alone.
%In total, we found 189\% as many bugs by using nondeterministic data-race candidates as PPs.

>>>>

{\bf Finding the same bugs faster.}
To show that Iterative Deepening is effective even for MC domains without data races, %such as message-passing distributed systems,
we ran the test suite with \quicksand~configured to explore only subsets of the hard-coded mutex PPs
(i.e., ignoring all data-race candidates).
%\footnote{
%Because \quicksand~is not yet instrumented to subset hard-coded PPs beyond the 4 ways shown in Figure~\ref{fig:id},
%we ran these tests for 2.5 hours on 4 CPUs each.
%Future work could parallelize QS-no-DR-PPs further; see \sect{\ref{sec:future}}.}.
%
The line QS-no-DR-PPs represents this experiment.
Even though SSS-MC mostly catches up to it by the end of the 10-hour budget,
QS-no-DR-PPs finds many more of the bugs much sooner.
Hence, for smaller arbitrary CPU budgets,
Iterative Deepening is likely to find bugs SSS-MC will miss.
%and for more ambitiously-sized tests,
%programmers can be more confident in the verification provided when \quicksand~times out with no bug found.
%We conclude that for smaller arbitrary CPU budgets, especially less than 1 hour,
%Iterative Deepening is likely to find bugs SSS-MC will miss.
%Moreover, it is also easy to imagine scaling up the size of each test case to test ,
%using more threads or longer sequences of API calls.
%We hope that is compelling even to users willing to spend many CPU-hours on testing.
%These results show explicitly that for arbitrary CPU-time budgets

@@

{\bf Finding new data-race bugs.}
The line QS-DRs represents \quicksand's bug-finding performance
with data-race PPs enabled.
We quickly pull ahead of SSS-MC, and ultimately conclude with 179\% as many bugs in total.
The break-even point is at a negligible 90 seconds.

@@

\begin{table*}[t]
	\begin{center}
		\small
	\begin{tabular}{r|c||c|c|c||c|c|c|c|c}
		% TODO: Make the "bugs" left half of this graph into a bar graph.
		& {\bf Num.} & {\bf SSS-MC} & {\bf QS} & {\bf DR-only} & {\bf Total} & {\bf Untested} & {\bf Malloc-} & {\bf Nondet.} & {\bf Nondet.} \\
		{\bf Test name} & {\bf tested} & {\bf bugs} & {\bf bugs} & {\bf bugs} & {\bf DR PPs} & {\bf DR PPs} & {\bf recycle DRs} & {\bf DR PPs} & {\bf DR bugs} \\
		\hline
		{\bct		} & 79	& 7	& 11	& 4	&	&	& 65	&	& 4	\\
		{\tej		} & 79	& 14 	& 25	& 11	&	&	& 471	&	& 6	\\
		{\mxtest	} & 79	& 1	& 10	& 9	&	&	& 7	&	& 1	\\
		{\paradise	} & 79	& 10	& 17	& 7	&	&	& 268	&	& 3	\\
		{\paraguay	} & 79	& 4	& 12	& 9	&	&	& 188	&	& 0	\\
		{\rwldgr	} & 79	& 26	& 30	& 3	&	&	& 157	&	& 3	\\
		\hline
		{\prisema	} & 59	&	&	&	&	&	& 0	&	&	\\
		{\alarmsimul	} & 44	&	&	&	&	&	& 37	&	&	\\
		{\waitsimple	} & 52	&	&	&	&	&	& 57	&	&	\\
		\hline
		{\bf Total}	& 629	&	&	&	&	&	& 1250	&	&
	\end{tabular}
	\end{center}
	\caption{Summary of number and types of bugs found by each test program.
		SSS-MC is the control; QS is \quicksand.
		``DR-only'' counts among \quicksand's bugs how many required data-race PPs to expose (\sect{\ref{sec:eval-sssmc}}).
	``Nondeterministic DRs'' counts among the DR-only bugs how many candidates required MC integration to identify (\sect{\ref{sec:eval-falseneg}}).
		TODO: more explanation}
	\label{tab:drbugs}
\end{table*}

%Figure~\ref{fig:bargraph}
% TODO: deal with the table
The left half of Table~\ref{tab:drbugs}
breaks down the number and types of bugs found by each test program.
% The astute observer will note that the sum of all types of failures observed is more than the total number of data-race bugs...
%TODO CAMREADY: Run a mutex expt where "all atomic instrs" are PPs. See how many bugs are missed anyway.
% Wwe re-ran the \mxtest control experiment with \landslide~hard-coded to preempt on any atomic instruction
% (as well as on the mutex API boundaries).
% Still, this smarter configuration for SSS-MC found only 99999999999 bugs of \quicksand's 13.
In \mxtest~in particular, the control experiment found dramatically fewer bugs (just 1),
even compared to the other test programs\footnote{
	The one bug SSS-MC found was in a fully-assembly lock implementation. {\tt yield()}'s return value clobbered a value stored in {\tt \%eax}, which could lead to a failure after two repeated contentions. Preempting only on {\tt yield()} (in the contention loop) was sufficient.}.
%Intuitively, this is due to our control experiment being able to preempt only on the boundaries of the API which
%Though for many applications of MC, assuming a correct lock implementation is sufficient,
Though it suffices for many applications of MC to assume correctly-implemented locks,
we consider this strong evidence that low-level synchronization code must be verified with data-race PPs.

@@

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{nondet-drs-1-v2.pdf}
	\caption{Some data-race candidates may not be identified during a single program execution.
		Using nondeterministic data races as PPs,
		\quicksand~found 189\% as many data-race bugs compared to using single-pass candidates alone.
	}
	\label{fig:dr-falsenegs}
\end{figure}
Figure~\ref{fig:dr-falsenegs} compares the types of data-race candidates necessary to expose each data-race bug in our test suite.
The first series represents the bugs found using PPs from single-pass data-race candidates,
% TODO: re read racefuzzer/portend paper to confirm this is true.
i.e., the state-of-the-art approach used by \cite{racefuzzer,portend}.
The second series shows all data-race bugs \quicksand~found,
which includes the former type as well as new bugs involving nondeterministic data-races.
\quicksand~found 68 data-race bugs in total, only 36 of which could be found with single-pass data-race candidates alone.
%In total, we found 189\% as many bugs by using nondeterministic data-race candidates as PPs.

### 2015-11-17 pldi-16/soundness.tex
minor rephrasing

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{struct x \{ int foo; int baz; \} *x;}} \\
	& \multicolumn{2}{c}{\texttt{struct y \{ int bar; \} *y;~~~~~~~~~~}} \\
	\\
	& Thread 1 & Thread 2 \\
	1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	3 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	5 & & \texttt{\hilight{commentblue}{// ...initialize...}}\\
	6 & & \texttt{publish(y);} \\
	7 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{A common execution pattern with {\tt malloc()} that produces false positive data race candidates.}
\label{fig:recycle}
\end{figure}
\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& Thread 1 & Thread 2 \\
	1 & \texttt{publish(x);} & \\
	2 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	3 & \texttt{\hilight{olivegreen}{free}(x);} \\
	4 & & \texttt{x2 = get\_published\_x();} \\
	5 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	6 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	7 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
\end{tabular}
\caption{If a single-pass data race detector discarded candidates matching the malloc-recycle pattern,
it would miss the bug in this adversarial program.}
\label{fig:recycle-bug}
\end{figure}

>>>>

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{struct x \{ int foo; int baz; \} *x;}} \\
	& \multicolumn{2}{c}{\texttt{struct y \{ int bar; \} *y;~~~~~~~~~~}} \\
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	3 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	5 & & \texttt{\hilight{commentblue}{// ...initialize...}}\\
	6 & & \texttt{publish(y);} \\
	7 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{A common execution pattern with {\tt malloc()} that produces false positive data race candidates.}
\label{fig:recycle}
\end{figure}
\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{publish(x);} & \\
	2 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	3 & \texttt{\hilight{olivegreen}{free}(x);} \\
	4 & & \texttt{x2 = get\_published\_x();} \\
	5 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	6 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	7 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
\end{tabular}
\caption{If a single-pass data race detector discarded candidates matching the malloc-recycle pattern,
it would miss the bug in this adversarial program.}
\label{fig:recycle-bug}
\end{figure}
